Bounds on the approximation of Kolmogorov complexity

### 2013-04-23 Bastards/src/bastards.tex
...

@@

\begin{lma}[{\cite{li1993introduction}[p 120]}]
For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^-k$. 
\end{lma}

@@

(Note that unlike the finite set case, here we do not make the complexity term conditional on the model)

@@

\begin{lma}
The probability of all strings with randomness deficiency above $k$ is less than $2^-k$
\end{lma}

@@

\begin{exm}
What is the probability that we sample a string $x$ with a randomness deficiency of at least half its length? In other words, if we use the optimal description, we gain $|x|/2$ over describing it in terms of $p_q$.
\end{exm}

@@

@@

\subsection*{}

\section*{Inference from bounded distributions}
\nocite{*}
\bibliographystyle{siam}
\bibliography{bastards}

>>>>

\begin{dfn}
The \emph{resource bounded Kolmogoroc complexity} of a string $x$ is defined as
\begin{align*}
K_U^t(x) = \min{|p| : U^t(p) = x}
\end{align*}
Where $t$ is a time-constructible fuction $t(n)$ and $U^t$ is a prefix Turing machine that only halts within $t(|x|)$ timesteps.
\end{dfn}

It is important to note that the invariance of the kolmogorov complexity under resource bounds is weaker than the unbounded version. Specifically, if we have some prefix turing machine $q$ which computes $x$ within bounds for an input of length $K^t_q(x)$, then we can simulate that Turing machine for a constant cost as normal, but only with a multiplicative constant penalty and an additive logarithmic one:
\[
K^{c \cdot t + \log t} \leq K^t_q(x)
\] 
For the following, we assume that our resource classes are sufficiently crude that this distinction doesnt matter.\footnote{For a more complete treatment of this issue, see \cite{li1993introduction}[pages 532--536]. The form of the penalty may change, depending on the computational metaphor.}

@@

\begin{lma}[{\cite{li1993introduction}[p 120]}]
For any finite set $A$, the proportion of strings with randomness deficieny above $k$ is less than $2^-k$.

\label{lma:rd-finitesets}
\end{lma}

@@

(Note that unlike the finite set case, here we do not make the complexity term conditional on the model) [@! does this matter?]

@@

\begin{lma}
The probability of all strings with randomness deficiency above $k$ is less than $2^{-k}$
\label{lma:rd-distributions}
\end{lma}

@@

Note that if we see a finite set as a uniform distribution over its elements, this proof provides a more general version of lemma \ref{lma:rd-finitesets}

@@

\begin{dfn}
Let $u$ be the prefix-free universal Turing machine. $p_u(x)$ is a \emph{universal distribution}. It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
\[
p_q(x) \leq c_q p_u(x)
\]
\label{dfn:universal-distribution}
\end{dfn}

If we bound the resources of the universal Turing machine, we get a resource bounded universal distribution:

\begin{dfn}
Let $u^t$ be the prefix-free universal Turing machine bounded by $t$. $p_{u^t}(x)$ is a \emph{resource-bounded universal distribution}. It dominates every computable distribution $q$ up to a multiplicative constant (dependent on $q$ but not $x$):
\[
p_q(x) \leq c_q p_{u^t}(x)
\]

\label{dfn:resource-bounded-universal-distribution}
\end{dfn}

Note again, that this only holds for a certain level of courseness in our resource-classes.

@@

\section*{Inference from bounded distributions}

\subsection*{Bastards}

We call a string a \emph{bastard} if resource-bounded Kolmogorov complexity shows it to be random, but under relaxation of these bounds, the string turns out to be compressible. This way, the probability of a bastard tells us the probability that we will make a mistake about the level of randomness in our data if we use a resource-bounded version of Kolmogorov complexity rather than the incomputable version.

\begin{dfn}
A \emph{$(t, k)$-bastard} is a string for which $K^t(x) - K(x) > k$
\end{dfn}

\begin{thm}
Let $q \in {\cal C}^t$. The probability of sampling a $(t, k)$-bastard from $q$ is negligable for large $k$.
\end{thm}
\begin{proof}

Let $B^t_k$ be the set of $(t, k)$-bastards. By definition \ref{dfn:resource-bounded-universal-distribution}, we have $p_q(B^t_k) \leq c_q p_{u^t}(B^t_k)$.

\begin{align*}
B^t_k &= \{x : K^t(x) - K(x) > k\} \\ 
	&\subseteq \{x : - \log p_{u^t}(x) - K(x) > k\} \\
	&= \{x : \delta(x | u^t) > k\}
\end{align*}

So the set of strings with randomness deficiency larger than $k$ for $u^t$ is a superset of the set of bastards. Since we sample string with such randomness deficiency with probability below $2^{-k}$, we have
\begin{align*}
p_{u^t}(B^t_k) < 2^{-k} 
\end{align*}
and 
\begin{align*}
p_q(B^t_k) < c_q \cdot 2^{-k} 
\end{align*}
\end{proof}

\begin{thm}
Let $q \in {\cal C}^u$ and $t < u$. The probability of sampling a $(t, k)$-bastard from $q$ is not negligable.
\end{thm}
\begin{proof}
\ldots
\end{proof}

\nocite{*}
\bibliographystyle{siam}
\bibliography{bastards}
